---
title: "data"
author: "Tyler Cobian"
date: "4/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Input the climate data
```{r}
clim <- read.table("clim.txt",  sep=" ", header=T)
```

Function

The basis of the model equations
Y = - 0.015 Tn2 - 0.0046 Tn2^2 - 0.07P1 _ 0.0043P1^2 + 0.28
Tn = minimum temperature in degrees celcius in that month
P = percipitation in mm in that month

1. Implement a model of almond yield anomaly (difference from average) in R based on this paper; You will be graded both on the “correctness” of the implementation and on “style”.  Make sure you add some error checking. Use coefficients in the paper as default values, but make sure you can vary them. 

```{r}

# I think this function works but we can probably come up with a more elegant solution 

almond_yield <- function( clim_data = clim,
                       a=0.015,
                       b=-0.0046,
                       c=-0.07,
                       d=0.0043,
                       e=0.28) {
  
  # Average monthly maximum daily temperature, and monthly precipitation from a data frame called clim  with columns year, month, precip and tmax_c

clim_month <-  clim_data %>%
group_by(month, year) %>%
summarize(meantmin = mean(tmin_c),
          meantmax = mean(tmax_c),
          precip=sum(precip))
  
   # Filter Jan and Feb data
  jan <- clim_month %>% 
    filter(month==1)
  
  feb <- clim_month %>% 
    filter(month==2)
  
  # Change column names
  colnames(jan) <- c("month", "year", "Tn", "Tm", "P")
  colnames(feb) <- c("month", "year", "Tn", "Tm", "P")
  
  # Data structure for yield annomalies
  yield_df <- data.frame(year = jan$year, YA = NA)
  
  # Loop through each year
  for (i in 1:length(yield_df$year)) {
   yield_df$YA[i] = a*feb$Tn[i] + b*(feb$Tn[i]^2) +c*jan$P[i] + d*(jan$P[i]^2) + e
}
  
  # Calculate max and min yields
  max_yield <- yield_df %>% 
    arrange(-abs(YA)) %>% 
    head(1)
  
  min_yield <- yield_df %>% 
    arrange(abs(YA)) %>% 
    head(1)
  
  # Create list with three elements
  yield_list <- list(yield_df, max_yield, min_yield)
  
  # Return list
return(yield_list)
  
}
  
almond_yield()
```




2. Have the function return almond yield anomaly for each year, and max and minimum yields over a time series of multiple year inputs





3. With the climate data provided (clim.txt), *use your model* to estimate how almond yield anomalies for each year in the record - graph the results.

```{r}
# Select fist element of the output list
clim_YA <- as.data.frame(almond_yield()[1])
clim_YA

# Create graph
graph_YA <- ggplot( clim_YA, aes( x = year, y = YA)) +
  geom_point() +
  xlab("Year") +
  ylab("Yield Anomaly") + # units?
  ggtitle("Almond Yield Anomalies (1989 - 2010)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

graph_YA
```

4. Use model to estimate the mean annual yield anomaly if the last 20 years had been 2C warmer
```{r}
# Make new clim data with min and max temp 2 degrees warmer
clim2 <- clim %>%
  mutate(tmin_c = tmin_c + 2) %>% 
  mutate(tmax_c = tmin_c + 2)

# Use function and convert to df
clim2_YA <- as.data.frame(almond_yield(clim_data = clim2)[1])

# Calculate mean
mean_clim2_YA <- mean(clim2_YA$YA)
mean_clim2_YA # 181.5714

```



5. Vary one of the parameter by sampling from a normal distribution with mean the value from Lobell et al., (2006) and standard deviation 10% of the mean -  create a box plot of  mean annual yield anomaly for baseline and 2C climate (where boxes show variation due to uncertanty in the parameter) 

```{r}

# Make a normal distribution for the tmin parameter in the model with the mean value of 4 degrees C which is the value used in the paper
tmin = rnorm(length(clim2), 12, 0.1)

# insert the new tmin value that is gained from a normal ditribution for tmin values in a new data frame clim3
clim3<- clim2 %>% 
  merge(tmin) %>% 
  select(-tmin) %>% 
  select(-y)

# run this new data through the function
clim_rnorm<- as.data.frame(almond_yield(clim_data = clim3)[1])

# Make a box plot with this new data
graph_rnorm<- ggplot(clim_rnorm, aes(x = year, y = YA, group = year))+
  geom_boxplot(aes(color = year))+
  geom_errorbar(mapping = aes(ymax = (YA*1.1), ymin = (YA*0.9)), position = position_dodge())
graph_rnorm

```

























