---
title: "data"
author: "Tyler Cobian"
date: "4/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Input the climate data
```{r}
clim <- read.table("clim.txt",  sep=" ", header=T)
```

Function

The basis of the model equations
Y = - 0.015 Tn2 - 0.0046 Tn2^2 - 0.07P1 _ 0.0043P1^2 + 0.28
Tn = minimum temperature in degrees celcius in that month
P = percipitation in mm in that month

1. Implement a model of almond yield anomaly (difference from average) in R based on this paper; You will be graded both on the “correctness” of the implementation and on “style”.  Make sure you add some error checking. Use coefficients in the paper as default values, but make sure you can vary them. 

2. Have the function return almond yield anomaly for each year, and max and minimum yields over a time series of multiple year inputs

```{r}
almond_yield <- function( clim_data = clim,
                       a=-0.015,
                       b=-0.0046,
                       c=-0.07,
                       d=0.0043,
                       e=0.28) {
  
  # Add in some error checking into the function
  
  # Make sure the climate data input is a dataframe
  if(class(clim_data) != "data.frame") return("Climate data input must be a data frame")
  
   # Make sure the climate data input contains the columns year, month, precip, tmin_c and tmax_c
  if(!all(has_name(clim_data, c("month", "year",  "month", "precip", "tmin_c", "tmax_c")))) return("Climate data input must contain the following columns: year, month, precip, tmin_c and tmax_c")
  
  # Make sure that the input for precipitation is larger than 0
  clim_data$precip = ifelse(clim_data$precip < 0, return("Input for precipitation must be a value larger than 0"), clim_data$precip)
  
  # Make sure that the maximum tempertaure will be larger than the minimum temperature
  clim_data$tmin_c = ifelse(clim_data$tmin_c > clim_data$tmax_c, return("Input for maximum temperature must be larger than input for minimum temperature"), clim_data$tmin_c)
  
  # Average monthly maximum daily temperature, and monthly precipitation from a data frame called clim  with columns year, month, precip and tmax_c

clim_month <-  clim_data %>%
group_by(month, year) %>%
summarize(meantmin = mean(tmin_c),
          meantmax = mean(tmax_c),
          precip=sum(precip))
  
   # Filter Jan and Feb data
  jan <- clim_month %>% 
    filter(month==1)
  
  feb <- clim_month %>% 
    filter(month==2)
  
  # Change column names
  colnames(jan) <- c("month", "year", "Tn", "Tm", "P")
  colnames(feb) <- c("month", "year", "Tn", "Tm", "P")
  
  # Data structure for yield annomalies
  yield_df <- data.frame(year = jan$year, YA = NA)
  
  # Loop through each year
  for (i in 1:length(yield_df$year)) {
   yield_df$YA[i] = a*feb$Tn[i] + b*(feb$Tn[i]^2) +c*jan$P[i] + d*(jan$P[i]^2) + e
}
  
  # Calculate max and min yields
  max_yield <- yield_df %>% 
    arrange(-abs(YA)) %>% 
    head(1)
  
  min_yield <- yield_df %>% 
    arrange(abs(YA)) %>% 
    head(1)
  
  # Change column names of max and min yields
  colnames(max_yield) <- c("Year", "Maximum Yield Anomaly")
  colnames(min_yield) <- c("Year", "Minimum Yield Anomaly")
  
  # Create list with three elements
  yield_list <- list(yield_df, max_yield, min_yield)
  
  # Return list
return(yield_list)
}
  
almond_yield()
```
3. With the climate data provided (clim.txt), *use your model* to estimate how almond yield anomalies for each year in the record - graph the results.

```{r}
# Select fist element of the output list
clim_YA <- as.data.frame(almond_yield()[1])
clim_YA

# Create graph
graph_YA <- ggplot( clim_YA, aes( x = year, y = YA)) +
  geom_point() +
  xlab("Year") +
  ylab("Yield Anomaly") + # units?
  ggtitle("Almond Yield Anomalies (1989 - 2010)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

graph_YA
```

4. Use model to estimate the mean annual yield anomaly if the last 20 years had been 2C warmer
```{r}
# Make new clim data with min and max temp 2 degrees warmer
clim2 <- clim %>%
  mutate(tmin_c = tmin_c + 2) %>% 
  mutate(tmax_c = tmin_c + 2)

# Use function and convert to df
clim2_YA <- as.data.frame(almond_yield(clim_data = clim2)[1])

# Calculate mean
mean_clim2_YA <- mean(clim2_YA$YA)
mean_clim2_YA # 181.2151

```

5. Vary one of the parameters by sampling from a normal distribution with mean the value from Lobell et al., (2006) and standard deviation 10% of the mean -  create a box plot of  mean annual yield anomaly for baseline and 2C climate (where boxes show variation due to uncertanty in the parameter) 

```{r}
# Make a normal distribution for the tmin parameter in the model with the mean value of 4 degrees C which is the value used in the paper
tmin = rnorm(length(clim2$year), 12, 0.1)

# insert the new tmin value that is gained from a normal ditribution for tmin values in a new data frame clim3
clim2$tmin_c<- tmin

# run this new data through the function
clim_rnorm<- as.data.frame(almond_yield(clim_data = clim2)[1])

# Make a box plot with this new data
graph_rnorm<- ggplot(clim_rnorm, aes(x = year, y = YA, group = year))+
  geom_col(aes(fill = year), show.legend = FALSE)+
  geom_errorbar(mapping = aes(ymax = (YA*1.1), ymin = (YA*0.9)), position = position_dodge())+
  labs(x = "Year", y = "Yield Anomoly", title = "Yearly Yield Anomoly With a Normal Distribution of Minimum Temperature")+
  theme_bw()
graph_rnorm

```

























